Abstract Data Type Inference

### 2015-03-25 patterns.tex
Fixed space hacks.

Computing $B$ thus amounts to computing its finite representation. In general,
this is achieved by enumerating the elements of $B$ while maintaining the
$\preceq$-minimal elements, and recognizing a condition under which all
elements of $B$ are related to the current set of
minimals~\cite{conf/lics/AbdullaCJT96, journals/tcs/FinkelS01}. In our case, we
recognize a condition on the weight-increasing enumeration of $B$ which
guarantees coverage by the current set of minimals. Formally, for all $i \in
\mathbb{N}$, we define
  \vspace{-1mm}
\begin{align*}
  B_i &= \set{ h \in B : \weight(h) \le i } \\
  B_i' &= \set{ h' \in B :
    \exists h \in B_i.\ h \preceq h' \text{ and } \weight(h') \le i+1
  }
    \vspace{-1mm}
\end{align*}
as, respectively, the histories of $B$ with at most $i$ matches and duplicates,
and those derived from $B_i$ with at most $i\!+\!1$ matches and duplicates. We
say that $A$ is \emph{predictable} if $B_i = B$ whenever $B_i' = B_{i+1}$. In
Section~\ref{sec:nature} we demonstrate that naturally-occurring ADTs are
predictable. Since the number of weight-$i$ histories is finite\footnote{As
noted in Section~\ref{sec:histories}, we consider equality between histories up
to renaming of operation identifiers.} for each $i \in \mathbb{N}$, each $B_i$
is computable.

@@

  The following four histories generate the complement of the atomic
  single-value register ADT, witnessing either an unmatched read operation:
  \vspace{-1mm}
\begin{verbatim}
  [1:X] read => 1 (RO)  #
\end{verbatim}
  \vspace{-1mm}
  a read of an uninitialized register occurring after some write:
\begin{verbatim}
  [1:1] write(1)            #
  [2:2] read => - (RO)    #
\end{verbatim}
    \vspace{-1mm}
  a read which happens before its matching write operation:
      \vspace{-1mm}
\begin{verbatim}
  [1:2] read => 1 (RO)  #
  [2:2] write(1)          #
\end{verbatim}
    \vspace{-1mm}
  and a read matching a write which is not the most recent:
      \vspace{-1mm}
\begin{verbatim}
  [1:1] write(1)        #
  [2:2] write(2)          #
  [3:1] read => 1 (RO)      #
\end{verbatim}
    \vspace{-1mm}
  Every sequential history not admitted by the atomic register ADT embeds at
  least one of these four histories.

>>>>

Computing $B$ thus amounts to computing its finite representation. In general,
this is achieved by enumerating the elements of $B$ while maintaining the
$\preceq$-minimal elements, and recognizing a condition under which all
elements of $B$ are related to the current set of
minimals~\cite{conf/lics/AbdullaCJT96, journals/tcs/FinkelS01}. In our case, we
recognize a condition on the weight-increasing enumeration of $B$ which
guarantees coverage by the current set of minimals. Formally, for all $i \in
\mathbb{N}$, we define
\begin{align*}
  B_i &= \set{ h \in B : \weight(h) \le i } \\
  B_i' &= \set{ h' \in B :
    \exists h \in B_i.\ h \preceq h' \text{ and } \weight(h') \le i+1
  }
\end{align*}
as, respectively, the histories of $B$ with at most $i$ matches and duplicates,
and those derived from $B_i$ with at most $i\!+\!1$ matches and duplicates. We
say that $A$ is \emph{predictable} if $B_i = B$ whenever $B_i' = B_{i+1}$. In
Section~\ref{sec:nature} we demonstrate that naturally-occurring ADTs are
predictable. Since the number of weight-$i$ histories is finite\footnote{As
noted in Section~\ref{sec:histories}, we consider equality between histories up
to renaming of operation identifiers.} for each $i \in \mathbb{N}$, each $B_i$
is computable.

@@

  The following four histories generate the complement of the atomic
  single-value register ADT, witnessing either an unmatched read operation:
\begin{verbatim}
  [1:X] read => 1 (RO)  #
\end{verbatim}
  a read of an uninitialized register occurring after some write:
\begin{verbatim}
  [1:1] write(1)            #
  [2:2] read => - (RO)    #
\end{verbatim}
  a read which happens before its matching write operation:
\begin{verbatim}
  [1:2] read => 1 (RO)  #
  [2:2] write(1)          #
\end{verbatim}
  and a read matching a write which is not the most recent:
\begin{verbatim}
  [1:1] write(1)        #
  [2:2] write(2)          #
  [3:1] read => 1 (RO)      #
\end{verbatim}
  Every sequential history not admitted by the atomic register ADT embeds at
  least one of these four histories.
